---
title: "MED64 Cross-Correlation Heatmap Analysis"
author: "Walter German"
output: html_notebook
---

This is notebook of cross-correlation analysis of MED-64 data

```{r imports_chunk, echo=FALSE}
library(tidyverse)
library(mema)
library(gplots)
library(functional)
```

### Dataset Filenames
```{r}
data_dir <- "/Users/walter/Data/margolis/HB_139_DAMGO_10-15/825_12h24m37s"
units_fname = "20200825_12h24m37s_units_ts.mat"
treatment = "20200825_12h24m37s_treatments_col_tx.csv"
experiment_tag = "825_12h24m37s"
save_path = "/Users/walter/Data/margolis/r_output"
```


### Load Dataset
```{r load_dataset}
# load the data
setwd(data_dir)

experiment <- mema::load_experiment(
  units_fname = units_fname,
  treatment = treatment,
  experiment_tag = experiment_tag,
  save_path = save_path)
```

#### Define the functions for histograms 
```{r}
unit_tx_activity <- function(df, unit, tx) {
  unit_tx = experiment[["firing"]] %>%
    filter(neuron_index == unit & treatment == tx)
  unit_tx[["time_step"]]
}

fr_bins <- function(df, unit, tx, brks) {
  unit_fr <-  unit_tx_activity(df, unit, tx)
  bins_fr <- hist(unit_fr, breaks = brks, plot = FALSE)
  bins_fr$count
} 
```


#### Creat bins for correlations
```{r}
units <- c(unique(experiment$firing$neuron_index))
brks = seq(0,600,0.1)  # histogram bins
fr_bins_bk <- Curry(fr_bins, 
                    brks = brks)
tx <- 'Start Baseline'

## This line takes a while, linear on limit
units_fr_bins <-units %>%
  map(invisible(fr_bins_bk), df = experiment, tx = tx)
```


#### Create Dataframe of Firing Rate Bins
```{r}
# create an empty dataframe
# This has index columns of x y unit numbers. 
# and then value columns for each pairwise unit xcorr
# thus rows equal num_units * num_units
xcorr_df <- data.frame(unitx = integer(), 
                       unity= integer(),
                       lag= double(), 
                       acf= double())
  

for (x in units) {
  for (y in units) {
      ccf_out = ccf(x = units_fr_bins[[x]], 
                    y = units_fr_bins[[y]], 
                    lag.max = 40,
                    plot = FALSE)
      ccf_rows = data.frame(unitx = x,
                            unity = y,
                            lag = ccf_out$lag,
                            acf = ccf_out$acf)
      xcorr_df <- rbind(xcorr_df, ccf_rows)
  }
}
```



```{r}
# zero lag xcorr only
xcorr_z <- xcorr_df %>%
  filter(lag == 0) %>%
  subset(select = -c(lag ))
```


```{r}
# order by unit to inspect 
xcorr_z_order <- xcorr_z[order(xcorr_z$unitx, xcorr_z$unity),]

glimpse(xcorr_z_order)
```


```{r}
# create wide dataframe for heatmap2 input
xcorr_z_wide <- xcorr_z %>%
  pivot_wider(names_from = unitx, values_from = acf )
```



```{r, fig.height=3, fig.width=3}
# primary purpose of running heatmap2 is to access its automatic ordering
# of the units to create aesthetic display. Better results than seriate.
plot_heatmap2 <- xcorr_z_wide %>%
  dplyr::select(-unity) %>%
  as.matrix() %>%
  heatmap.2(dendrogram="none",
            key=FALSE,
            trace = 'none',
            main = "zero lag correlogram",
            xlab = "unit number",)

heatmap_order <- plot_heatmap2$rowInd
```


```{r}
# Create four dataframes. Each has dimensions that are needed as inputs 
# depending on the type of final plot desired. Character labels for units
# allows for flexibility in axis tick labels.
# df_long  # labels as numeric and units labeled correctly
# df_long_char  # labels as character and units labeled correctly.
#
# df_long_order   # labels as numeric and renumbered so "unit 1" is renamed to
#                 # where heatmap2 order looks best
# df_long_order_char  # labels as character and renumbered


# default df names are character, even if passed numeric
unit_names <- units # as.character(units) 

df_wide_corr_z <- data.frame(xcorr_z_wide %>%
  dplyr::select(-unity) %>%
  as.matrix())

colnames(df_wide_corr_z) <- unit_names
rownames(df_wide_corr_z) <- colnames(df_wide_corr_z)

reorder <- function(df, new_col_order) {
  df[new_col_order]
}

rename_cols <- function(df, new_col_names) {
  colnames(df) <- new_col_names
  df
}

col_to_num <- function(df, col_name) {
  df[col_name] <- as.numeric(as.matrix(df[col_name]))
  df
}

col_to_char <- function(df, col_name) {
  df[col_name] <- as.character(as.matrix(df[col_name]))
  df
}


df_wide <- df_wide_corr_z %>%
  add_column(unitx = unit_names, .before = 1)

df_long <- df_wide %>%
  pivot_longer(cols = as.character(units), names_to = "unity", values_to = "acf") %>%
  col_to_num("unity")

df_long_char <- df_long %>%
  col_to_char("unitx") %>%
  col_to_char("unity")
  
df_wide_order <- df_wide_corr_z %>%
  reorder(heatmap_order) %>%
  t() %>%
  data.frame() %>%
  rename_cols(unit_names) %>%
  reorder(heatmap_order)  %>%
  add_column(unitx = unit_names, .before = 1) %>%
  rename_cols(c("unitx", unit_names))

df_long_order <- df_wide_order %>%
  pivot_longer(cols = as.character(units), names_to = "unity", values_to = "acf") %>%
  col_to_num("unity")

df_long_order_char <- df_long_order %>%
  col_to_char("unitx") %>%
  col_to_char("unity")
```



```{r}
# global changes to saved plot formatting
theme_update(plot.title = element_text(hjust = 0.5))
```


### Heatmap with units labeled with correct axis tick order
[heatmap example code](https://r-charts.com/correlation/heat-map-ggplot2/)

```{r}
# https://r-charts.com/correlation/heat-map-ggplot2/
# replace xcorr_z with df_long_order or df_long_char

plot <- ggplot2::ggplot(df_long_char, 
                        aes(x = unitx, y = unity, fill = acf)) +
  geom_tile(color = "white",
            lwd = 0.1,
            linetype = 1) +
  scale_x_discrete(name ="Unit X", 
                    limits=heatmap_order_str) +
  scale_y_discrete(name ="Unit Y", 
                    limits=heatmap_order_str) +
  ggplot2::scale_fill_viridis_c(
    "Correlation",
    limits = c(0, 1),
    breaks = c(0, .25, .5, .75, 1),
    expand = c(0, 0)) +
  # scale_fill_gradient(low = "grey", high = "red")+
  coord_fixed() +
  ggplot2::guides(fill = ggplot2::guide_colourbar(nbin = 100, barwidth = 10, barheight = 1)) +
  #guides(fill = guide_colourbar(barwidth = 20,barheight = 0.8, title = "correlation")) +
  ggtitle("Unit Cross-Correlation Heatmap") +
  theme(plot.title = element_text(hjust = 0.5)) +
  #ggplot2::scale_x_discrete("") +
  #ggplot2::scale_y_discrete("") +
  ggplot2::theme(legend.position = "bottom",
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot2::ggsave(
  filename = "../product/cross_correlation_2022-01-04_tick-order.pdf",
  plot = plot,
  width = 8,
  height = 9)

plot
```


## Analyse the high correlation pairs
Zoom in onto the highly correlated units to look for signs of mislabeled events
 
```{r}
# zero lag xcorr only
xcorr_z <- xcorr_df %>%
  filter(lag == 0) %>%
  subset(select = -c(lag ))

# order by acf value to find large xcorr events
xcorr_z_max <- xcorr_z[order(desc(xcorr_z$acf)),] %>%
  filter(unitx < unity)

# Show where cut off is for the xcorr
plot(xcorr_z_max$acf,
     main = "Distribution of cross-correlation values at zero lag",
     xlab = "ordered by unit*unit x-corr value",
     ylab = "Cross-Correlation")
```



```{r}
# high x-correlation cutoff value 
cutoff <- 3

# order by acf value to find large xcorr events
xcorr_z_max_cut <- xcorr_z_max[1:cutoff,]

hi_x_units <- unique(append(xcorr_z_max_cut$unitx, xcorr_z_max_cut$unity))

# also take those that are anti-correlated acf < 0 
xcorr_df_sub <- xcorr_df %>%
  filter(unitx %in% hi_x_units) %>%
  filter(unity %in% hi_x_units)
```


```{r}
# plot the grid of select xcorrs
ggplot(xcorr_df_sub, aes(lag, acf)) + 
  geom_bar(stat="identity") +
  facet_grid(unitx ~ unity) +
  ggtitle("Cross-correlograms of highest cross-correlation units") +
  labs(x = "lag (100ms)")
```

### Display Units with high Firing Rate
Display Units with high Firing Rate to check that they are also
high cross-correlation

```{r}
units_fr <- units %>%
  map(fr_bins, df = experiment, tx = tx, brks = brks) %>%
  map(mean) %>%
  unlist()

df_firing_rate <- data.frame(unit=units, firing_rate=units_fr)

head(df_firing_rate[order(desc(df_firing_rate$firing_rate)),], 10)
```



